Proposed Solution

### 2017-06-26 ISSTA17/secs/completeConfigC.tex
submitted version
camera ready - take 1

ConfigC is complete but unsound.
We first define the set of incorrect files based on ground truth $Inc_{gt}$, and the set of files which are predicted by the model to be incorrect, $Inc_{prd}$.
To define $Inc_{gt}$, we will introduce the set of relations that are necessary and breaking based on ground truth, $Nec_{gt}$ and $Br_{gt}$ respectively.
The definition of $Inc_{prd}$ follows from the description of the algorithm.

\begin{align}
Inc_{gt} = \{&\forall C, \exists r \in Nec_{gt}, r \notin M(C)\\ \nonumber
 &\lor \exists r \in Br_{gt}, r \in M(C)\} \\
Inc_{prd} = \{&\forall C, \exists r \in Nec, r \notin M(C)\}
\end{align}

Our definition of sound and complete are as follows:

\begin{align}
&\forall C \in Inc_{gt}, C \in Inc_{prd} \text{  Complete} \\
&\not \exists C \in Inc_{prd}, C \notin Inc_{gt} \text{  Sound}
\end{align}

Completeness follows from showing that $Inc_{gt} \subseteq Inc_{prd}$.
The algorithm is complete if, for any file which breaks a relation for which we have a template (a misconfiguration) and have seen an correct example, the system will mark that file as incorrect.
Since the model always maintains the strongest conditions for correctness for any file, the algorithm is clearly complete.
Completeness corresponds to no false negatives.
However, this algorithm is unsound, as some correct file may be marked as incorrect, i.e. ConfigC returns a false positive.
We have a high false positive rate since the model is an over-approximation of the true model.

>>>>

\subsection{Completeness}

We demonstrate that ConfigC is complete but unsound.
That is, we will always detect a misconfiguration, but may also report correct files as misconfigurations.
We first define the set of incorrect files based on ground truth $Inc_{gt}$, and the set of files which are predicted by the model to be incorrect, $Inc_{prd}$.
To define $Inc_{gt}$, we will introduce the set of relations that are necessary and breaking based on ground truth, $Nec_{gt}$ and $Br_{gt}$ respectively.
The definition of $Inc_{prd}$ follows from the description of the algorithm.
%
\begin{align}
Inc_{gt} = \{&\forall C, \exists r \in Nec_{gt}, r \notin M(C)\\ \nonumber
 &\lor \exists r \in Br_{gt}, r \in M(C)\} \\
Inc_{prd} = \{&\forall C, \exists r \in Nec, r \notin M(C)\}
\end{align}

Our definition of sound and complete are as follows:
%
\begin{align}
&\forall C \in Inc_{gt}, C \in Inc_{prd} \text{  Complete} \\
&\not \exists C \in Inc_{prd}, C \notin Inc_{gt} \text{  Sound}
\end{align}

Completeness follows from showing that $Inc_{gt} \subseteq Inc_{prd}$.
The algorithm is complete if, for any file which breaks a relation for which we have a template (a misconfiguration) and have seen a correct example, the system will mark that file as incorrect.
Since the model always maintains the strongest conditions for correctness for any file, the algorithm is clearly complete.
Completeness corresponds to no false negatives.
However, this algorithm is unsound, as some correct file may be marked as incorrect, i.e. ConfigC returns a false positive.
We have a high false positive rate since the model is an over-approximation of the true model.

### 2017-06-26 ISSTA17/secs/intro.tex
camera ready - take 1

%What have we already done (labeled correct)
In our previous work~\cite{Santolucito2016} we built the tool ConfigC to verify configuration files from a set of correct examples.
ConfigC takes as a training set a set of correct configuration files and builds a set of rules describing a language model.
This model is then used to check if a new configuration file adheres to those rules.
If a file is incorrect, the tool identifies the location of error and reports what is incorrect.
For example, ConfigC might output \texttt{Ordering Error: Expected "extension mysql.so"} before \texttt{"extension recode.so"}.
While ConfigC's results are promising, the main weakness is that the learning can only use correct configuration files.
Additionally, there is a relatively high false positive rate (marking correct files as incorrect).

%What do we 'plan' to do to extend this (labeled correct and incorrect)
Here we propose to extend that work, which only works on correct examples, to work on both correct and incorrect examples by building an SMT formula to find a classification model.
By learning over both correct and incorrect examples, our hypothesis is that we will be able to decrease the high false positive rate observed in ConfigC, making the tool more useful to end users.
In order to extend ConfigC, we first provide a new description of ConfigC in terms of version space learning.
This description allows us to understand how the error reports are are produced in ConfigC.
Any extension to ConfigC must maintain the \textit{legibility} of learning - the tool must not only classify configuration files as correct or incorrect, but also identify the contributing factors (misconfigurations) to that classification.

%legibility issue
Legibility is a difficult property to achieve in machine learning and an active area of research~\cite{lei2016rationalizing}.
Although machine learning has been used in various software analysis techniques, such as programming-by-example \cite{lau2000version} and invariant synthesis \cite{garg2014ice}, this requires careful selection of the algorithms.
<any popular machine learning algorithms, such as neural nets and n-gram models, notoriously lack legibility and so are not well-suited to to provide simple justifications for their classification results.
The lack of justification for the results limits the applicability to software analysis, where providing feedback to a user is critical.
For example, in the case of error detection, the system should not only report which files have errors, but also locate the errors so a user may fix them.
By using non-probabilistic version space learning, we expect to have a highly level of legibility in our learning results.

%correctness issue
Additionally, probabilistic approaches to machine learning cannot be guaranteed to be complete, that is they are not guaranteed to always find an error if one exists.
We use version space learning to address the need for a formal completeness guarantee in an automated model generation for verification.
Version space learning is a technique for logical constraint classification~\cite{mitchell82}.
As a logical constraint system, version space learning not only provides justification of its classification results, but is has the foundation to build a completeness proof without the need for probabilistic thresholds.

@@

%evaluation plans
To test this approach in practice, we plan to implement our algorithm to check for TravisCI\footnote{\url{http://www.travis-ci.com}} configuration errors.
TravisCI is a continuous integration tool connected to Github that allows programmers to automatically run their test suite on every code update (commit).
A user adds a configuration file to the repository that enables TravisCI and specifies build conditions, such as which compiler to use, which dependencies are required, and a set of benchmarks to test.
This ensures the tool can always be automatically built correctly on a fresh machine.
The algorithm will utilize the temporal ordering on code in Github histories to allow us to detect potential build errors without actually building, saving valuable programmer and server time.

@@

>>>>

%What have we already done (labeled correct)
In our previous work~\cite{Santolucito2016}, we built the tool ConfigC to verify configuration files from a set of correct examples.
ConfigC takes as a training set a set of correct configuration files and builds a set of rules describing a language model.
This model is then used to check if a new configuration file adheres to those rules.
If a file is incorrect, the tool identifies the location of error and reports what is incorrect.
For example, when analysing MySQL configurations, ConfigC might output \texttt{Ordering Error: Expected "extension mysql.so"} before \texttt{"extension recode.so"}.
While ConfigC's results are promising, the main weakness is that the learning can only use correct configuration files.
Additionally, there is a relatively high false positive rate (marking correct files as incorrect).

%What do we 'plan' to do to extend this (labeled correct and incorrect)
Here we propose to extend that work, which only works on correct examples, to work on both correct and incorrect examples by building an SMT formula to find a classification model.
By learning over both correct and incorrect examples, our hypothesis is that we will be able to decrease the high false positive rate observed in ConfigC, making the tool more useful to end users.
In order to extend ConfigC, we first provide a new description of ConfigC in terms of version space learning~\cite{mitchell82}.
This description allows us to understand how the error reports are produced in ConfigC.
Any extension to ConfigC must maintain the \textit{legibility} of learning - the tool must not only classify configuration files as correct or incorrect, but also identify the contributing factors (misconfigurations) to that classification.

%legibility issue
Legibility is a difficult property to achieve in machine learning and an active area of research~\cite{lei2016rationalizing}.
Although machine learning has been used in various software analysis techniques, such as programming-by-example \cite{lau2000version} and invariant synthesis \cite{garg2014ice}, this requires careful selection of the algorithms.
Many popular machine learning algorithms, such as neural nets and n-gram models, notoriously lack legibility and so are not well-suited to to provide simple justifications for their classification results.
The lack of justification for the results limits the applicability to software analysis, where providing feedback to a user is critical.
For example, in the case of error detection, the system should not only report which files have errors, but also locate the errors so a user may fix them.
By using non-probabilistic version space learning, we expect to have a highly level of legibility in our learning results.

%correctness issue
Additionally, probabilistic approaches to machine learning cannot be guaranteed to be complete, that is they are not guaranteed to always find an error if one exists.
We use version space learning to address the need for a formal completeness guarantee in an automated model generation for verification.
Version space learning is a technique for logical constraint classification.
As a logical constraint system, version space learning not only provides justification of its classification results, but is has the foundation to build a completeness proof without the need for probabilistic thresholds.

@@

%evaluation plans
To test this approach in practice, we plan to implement our algorithm to check for TravisCI\footnote{\url{http://www.travis-ci.com}} configuration errors.
TravisCI is a continuous integration tool connected to Github that allows programmers to automatically run their test suite on every code update (commit).
A user adds a configuration file to the repository that enables TravisCI and specifies build conditions, such as which compiler to use, which dependencies are required, and a set of benchmarks to test.
This gives the developer confidence the code can always be automatically built correctly on a fresh machine.
The algorithm will utilize the temporal ordering on code in Github histories to allow us to detect potential build errors without actually building, saving valuable programmer and server time.

@@

While ConfigC was applied to MySQL configurations, and we propose our next application as TravisCI, this approach can be used for any configuration language.
TravisCI is just one example of an easily accessible set of labeled configuration files with a partial order.

### 2017-06-26 ISSTA17/secs/limits.tex
camera ready - take 1

We have made two assumptions in Eq. \ref{eq:E1}; first that the summary $P_t$ sufficiently detailed, i.e. contains every piece of information that might lead to a build error, and second that the model $M(P_t)$ will learn all relations that might lead to a build error, i.e. has a template for all types of errors.
While these are the strong assumptions - our algorithm is able to detect cases where these assumptions are not met.
If we cannot find a solution for Eq. \ref{eq:PE}, it means either $P_t$ or $M(P_t)$ has been underspecified and is insufficiently detailed.
It is then the user's responsibility to expand the definitions accordingly.

In addition to the sufficient detail of $P_t$ and $M(P_t)$, we must pick a trusted base.
Since it is possible for TravisCI to have bugs, it is possible that between versions of TravisCI, two identical program summaries may have different build statuses.
Since version space learning is (generally) intolerant of noise, we require that
$P_t = P_{t'} \implies S(P_t) = S(P_{t'})$.
If we allow for noisy data, for example in the case that it is not possible to identify a trusted base, this predicate will no longer hold.
This situation would require a noise tolerant version space learning~\cite{hong1997generalized}, though this may degrade the completeness guarantees.

>>>>

We have made two assumptions in Eq. \ref{eq:E1}; first that the summary $P_t$ sufficiently detailed, i.e. contains every piece of information that might lead to a build error, and second that the model $M(P_t)$ will learn all relations that might lead to a build error, i.e. has a template for all types of errors.
While these are strong assumptions, our algorithm is able to detect cases where these conditions are not met.
If we cannot find a solution for Eq. \ref{eq:PE}, it means either $P_t$ or $M(P_t)$ has been underspecified and is insufficiently detailed, and the definitions must be expanded accordingly.

In addition to the sufficient detail of $P_t$ and $M(P_t)$, we must pick a trusted base.
Since TravisCI may not always be backwards compatible, it is possible that between versions of TravisCI, two identical program summaries may have different build statuses.
Since version space learning is (generally) intolerant of noise, we require that
$P_t = P_{t'} \implies S(P_t) = S(P_{t'})$.
If we allow for noisy data, for example in the case that it is not possible to identify a trusted base, this predicate will no longer hold.
This situation would require a noise tolerant version space learning~\cite{hong1997generalized}, though this may degrade the completeness guarantees.

### 2017-06-26 ISSTA17/secs/related.tex
camera ready - take 1

Static configuration verification has been proposed as an effective way  
to tackle misconfiguration problems~\cite{xu15systems}.
Nevertheless, a practical and automatic configuration
verification approach still remains an open problem.

@@

One approach to preemptive error detection is to use predefined constraints, or correctness rules, for checking.
After encoding these constraints, it is straightforward to report errors if a target configuration file does not satisfy these rules.
For example, ConfValley is specification language for administrators to provide constraints on their systems' configurations~\cite{huang15confvalley}. 
This approach requires that administrators manually write specifications - an error-prone and difficult process. 
In contrast, by using a learning approach, our work does not require users to manually write any specifications.

@@

\para{White-box}
In our approach we assume a ``black-box'' model where we cannot access the system source code.
This is as opposed to a ``white-box'' model, where the source code is available and can leveraged in the configuration analysis.
Although this is a strong assumption, PCheck~\cite{xu16early} explores misconfiguration detection by adding configuration checking code directly in the system source code.
The system is then emulating with various potential configurations of the system.
One drawback to this approach is that for some systems (\eg, ZooKeeper) whose behavior is hard to emulate, PCheck cannot automatically generate the corresponding checking code.
Due to the emulation based testing strategy, PCheck's scope is limited to reliability problems caused by misconfiguration parameters. 
In contrast, \app is a ``black-box'' approach and only requires a training set of configuration files to learn rules.
By using the version space learning strategy of examples, we are also able to detect general misconfiguration issues that are outside the scope of emulation testing (\eg memory or thread usage settings), including performance, security, availability and reliability.

\iffalse
%TODO rewrite
\para{Misconfiguration diagnosis}
Misconfiguration diagnosis approaches have been proposed to address configuration problems post-mortem.
For example, ConfAid~\cite{attariyan10automating} 
and X-ray~\cite{attariyan12x-ray} use dynamic information
flow tracking to find possible configuration errors that may have resulted in
failures or performance problems. AutoBash~\cite{su07autobash} 
tracks causality and automatically fixes 
misconfigurations. Unlike \app, most misconfiguration
diagnosis efforts aim at finding errors after system
failures occur, which leads to prolonged recovery time.
\fi

>>>>

@@

Another language based approach is used by ConfValley, a specification language for administrators to provide constraints on their systems' configurations~\cite{huang15confvalley}. 
This approach requires that administrators manually write specifications - an error-prone and difficult process. 
In contrast, by using a learning approach, our work does not require users to manually write any specifications.

@@

\para{White-box}
In our approach we assume a ``black-box'' model where we cannot access the system source code.
This is as opposed to a ``white-box'' model, where the source code is available and can leveraged in the configuration analysis.
Although this is a strong assumption, PCheck~\cite{xu16early} explores misconfiguration detection by adding configuration checking code directly in the system source code.
The system is then emulating with various potential configurations of the system.
One drawback to this approach is that for some systems (e.g. ZooKeeper) whose behavior is hard to emulate, PCheck cannot automatically generate the corresponding checking code.
Due to the emulation based testing strategy, PCheck's scope is limited to reliability problems caused by misconfiguration parameters. 
In contrast, our work is a ``black-box'' approach and only requires a training set of configuration files to learn rules.
By using the version space learning strategy of examples, we are also able to detect general misconfiguration issues that are outside the scope of emulation testing (e.g. memory or thread usage settings), including performance, security, availability and reliability.

\iffalse
%TODO rewrite
\para{Misconfiguration diagnosis}
Misconfiguration diagnosis approaches have been proposed to address configuration problems post-mortem.
For example, ConfAid~\cite{attariyan10automating} 
and X-ray~\cite{attariyan12x-ray} use dynamic information
flow tracking to find possible configuration errors that may have resulted in
failures or performance problems. AutoBash~\cite{su07autobash} 
tracks causality and automatically fixes 
misconfigurations. Unlike our work, most misconfiguration
diagnosis efforts aim at finding errors after system
failures occur, which leads to prolonged recovery time.
\fi

### 2017-06-26 ISSTA17/secs/temporal.tex
camera ready - take 1

While in ConfigC, configuration files were labeled in isolation from their system by an expert, TravisCI configurations are automatically labeled with respect to their containing repository.
Because the correctness of TravisCI configuration files is dependent on the code it builds, we must model more than just the configuration file, $C$. 
We introduce a program summary $P_t$, which is a combination of $C$ as well as the information relevant to the learning process from the repository.
The subscript on $P_t$ is a timestamp tag based on the ordered git commit history.
In the case of TravisCI, this include the \verb|.travis.yml| file, as well as code features that may effect build status, such as programming language and a list of imported libraries.
The summary must be \textit{sufficiently detailed}, that is it must contain every piece of information that might lead to a build error.

%IMPORTANT, but no space
However, most versioning systems do not limit version history to a single linear timeline.
For example, Git features the ability to \textit{branch}, which allows to simultaneous commit chains.
To handle the start of a branch, add a superscript to indicate the branch, and restart the counter on a branch.
To handle the merge of two branches $P_{t}^{x}$ and $P_{t'}^{y}$, step to $P_{t+1}^{x}$, where $x$ is the mainline branch.
We then say that $P_{t'}^{y}$ has no successor commit $P_{t'+1}^{y}$.
Note the ability to handle branching implies this method is suitable to any training set which has a partial order.

@@

In contrast with ConfigC, we now consider both incorrect and correct and so must introduce the \textit{general boundary}.
The general boundary is the dual of the specific boundary, and is the most relaxed requirement for a positive classification.
We denoted specific boundary as the set of necessary relations $Nec$, and now denote the general boundary as the set of breaking relations $Br$.
With this notation, we can formally express the requirement that the program summary is sufficiently detailed.
\begin{align}
  \forall S(P_t)=Err, \exists r \in M(P_t), r \in Br \label{eq:E1}
\end{align}

@@

@@

While existential set operations can be expensive on large sets for an SMT solver, in our application this is not the case.
Thanks to the practice of making incremental commits when using source control, these sets will be small and the SMT will be fairly cheap.
In fact, the above implication generalizes to $P_{t,t+n}$, but for efficiency we must require that $M(P_{t})\ \setminus M(P_{t+n})$ is manageably small.
The definition of small here remains to be experimentally determined.

>>>>

While in ConfigC, configuration files were manually labeled in isolation from their system by an expert, TravisCI configurations are automatically labeled with respect to their containing repository.
Because the correctness of TravisCI configuration files is dependent on the code it builds, we must model more than just the configuration file, $C$. 
We introduce a program summary $P_t$, which is a combination of $C$ as well as the information relevant to the learning process from the repository.
The subscript on $P_t$ is a timestamp tag based on the ordered git commit history.
In the case of TravisCI, this include the \verb|.travis.yml| file, as well as code features that may affect build status, such as programming language and a list of imported libraries.
The summary must be \textit{sufficiently detailed}, that is, it must contain every piece of information that might lead to a build error.

%IMPORTANT, but no space
However, most versioning systems do not limit version history to a single linear timeline.
For example, Git features the ability to \textit{branch} and create temporally simultaneous commit chains.
To handle the start of a branch, we add a superscript to indicate the branch ($x$ and $y$ here).
To handle the merge of two branches $P_{t}^{x}$ and $P_{t'}^{y}$, step to $P_{t+1}^{x}$, where $x$ is the mainline branch.
We then say that $P_{t'}^{y}$ has no successor commit $P_{t'+1}^{y}$.
Note the ability to handle branching implies this method is suitable to any training set which has a partial order.

@@

In contrast with ConfigC, we now consider both incorrect and correct examples and so must introduce the \textit{general boundary}.
The general boundary is the dual of the specific boundary, and is the most relaxed requirement for a positive classification.
We denoted specific boundary as the set of necessary relations $Nec$, and now denote the general boundary as the set of breaking relations $Br$.
With this notation, we can formally express the requirement that the program summary is sufficiently detailed.
\begin{align}
  \forall S(P_t)=Err, \exists r \in M(P_t), r \in Br \label{eq:E1}
\end{align}

@@

In fact, to make full use of the partial order, we can generalize Eq \ref{eq:PE} to $P_{t,t+n}$.
For example, the pattern $Pass_0,Err_1,Pass_2,Err_3$ yields a PE pattern on $P_{0,1}$, $P_{2,3}$, and $P_{0,3}$. 
More complex chains could be built in the case of branching.

@@

While existential set operations can be expensive on large sets for an SMT solver, in our application this is not the case.
Thanks to the practice of making incremental commits when using source control, these sets will be small and the SMT will be fairly cheap.
However, for efficiency we must choose a maximum $n$ such that $M(P_{t})\ \setminus M(P_{t+n})$ is manageably small.
The definition of small and selection of maximum $n$ remains to be experimentally determined.

### 2017-06-26 ISSTA17/secs/vspace_verify.tex
submitted version
camera ready - take 1

Our definition of a configuration file, $C$, is a file that can be transformed to an ordered list of $(keyword, value)$ pairs, called $Line$s.
ConfigC builds a model for a single file, $M(C)$, that is the set of all possible relations between the lines in the file.
A relation is described by user provided templates.
A template is a function taking some number of lines and determining a relation on the keywords of the lines ($Line^{n} \rightarrow Rel_{k1,..,kn}$).
As an example, ConfigC might derive the relation \texttt{extension mysql.so} comes before \texttt{extension recode.so} by using the ordering template.
ConfigC features the following default templates.

\begin{center}
\begin{tabular} {|c |c |c |}
 \hline
 Template & Arity & Relation \\
\hline
 ordering &            2 & Before $\vert$ After $\vert$ None \\
\hline
 coarse relation &     2  & $< | == | >$  \\
\hline
 fine-grain relation & 3  & $< | == | >$  \\
\hline
 type &                1  & String $\vert$ Int $\vert$ Boolean $\vert$ IP  \\
\hline
 missing entry &       2 & Required $\vert$ Not \\
\hline
\end{tabular}
\end{center}

It is assumed if a file is correct, all relations in that file are in the set of necessary relations, $Nec$, for any other file to be correct ($Correct(C) \implies \forall r \in M(C), r \in Nec$).
In this way, the initial model is built by creating the strongest conditions for a correct file, called the \textit{specific boundary} in version space learning.
This model is then iteratively relaxed as more examples are seen, a process called \textit{candidate elimination}.
To relax the model, two sets of relations from two files are merged into a single consistent set.
\begin{lstlisting}
n = {}
for (c in files):
  n1 = M(c)
  n = merge(n, n1)
\end{lstlisting}

Since we maintain the strongest condition for correctness, this approach will identify many correct files as incorrect, which we call a high false positive rate.
This is a problem when a user is then asked to manually review many files for errors, when the files are in fact correct.

@@

>>>>

Our definition of a configuration file, $C$, is a file that can be transformed to an ordered list of $(keyword, value)$ pairs, called $Line$s.
ConfigC builds a model for a single file, $M(C)$, that is the set of all possible relations between the lines in the file.
A relation is described by user provided templates.
A template is a function taking some number of lines and determining a relation on the keywords of the lines ($Line^{n} \rightarrow Rel_{k1,..,kn}$).
As an example, ConfigC might derive the relation \texttt{extension mysql.so} comes before \texttt{extension recode.so} by using the ordering template.
ConfigC includes default templates for ordering, integer relations ($<,=,>$), type errors, and missing entries (two keywords must appear together irrespective of order).

It is assumed if a file is correct, all relations in that file are in the set of necessary relations, $Nec$, for any other file to be correct ($Correct(C) \implies \forall r \in M(C), r \in Nec$).
In this way, the initial model is built by creating the strongest conditions for a correct file, called the \textit{specific boundary} in version space learning.
This model is then iteratively relaxed as more examples are seen, a process called \textit{candidate elimination}.
To relax the model, two sets of relations from two files are merged into a single consistent set.

@@

Since we maintain the strongest condition for correctness, this approach will identify many correct files as incorrect, which we call a high false positive rate.
This is a problem when a user is then asked to manually review many files for errors, when the files are in fact correct.
However, the benefit to this approach is that, given the coverage of the training set, every misconfiguration will be detected.
